---
title: "Active & passive example: aggregate versus multilevel analysis"
author: "anonymous"
date: "03/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This analysis is a dataset analogous to Table 1, but with more variation added (otherwise terms become inestimable). The point of this analysis is to show how a summary analysis of Table 1 (binomial test, Chi-square test, Fisher's exact test etc.) yields a significant result that goes away once a multilevel is fitted that accounts for by-speaker variation.

## Prelims:

Load packages:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(brms)
```

Load mock data:

```{r, warning = FALSE, message = FALSE}
voice <- read_csv('../data/active_passive_example.csv')
```

## Faulty aggregate analysis

Aggregate over individuals:

```{r}
counts <- voice %>% select(active:passive) %>% 
  summarize(active = sum(active),
            passive = sum(passive)) %>% unlist() %>% as.vector()
```

Check counts:

```{r}
counts
```

Perform a binomial test:

```{r}
binom.test(counts)
```

## Generalized linear model accounting for by-speaker variation

Next we fit a GLM for which the data needs to be represented in a different way (in long format):

```{r}
voice_long <- voice %>%
  pivot_longer(active:passive,
               names_to = 'voice',
               values_to = 'count')
```

Check:

```{r}
voice_long
```

Fit a glm, in this case a Poisson model since the response variable in question is a count variable.

```{r}
# Run chains in parallel:
options(mc.cores = parallel::detectCores())

pois_mdl <- brm(count ~ voice + (1 + voice|speaker),
                data = voice_long,
                family = poisson,
                
                # Prior:
                prior = c(prior(normal(0, 1), class = b)),
                
                # MCMC control parameters:
                seed = 666,
                init = 0,
                chains = 4,
                warmup = 4000, iter = 6000,
                control = list(adapt_delta = 0.99,
                               max_treedepth = 13))
```

Check:

```{r}
summary(pois_mdl)
```

Not a lot of evidence for this effect anymore after speaker information has been incorporated.

Assess the posterior probability of the voice effect:

```{r}
posts <- posterior_samples(pois_mdl)$b_voicepassive
sum(posts > 0) / length(posts)
```

Again show that there is not a lot of evidence for the active vs. passive difference anymore.







