---
title: "Mixed model example"
author: "anonymous"
date: "07/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This script creates simulated data for Figure 1 and performs a mixed model analysis of this.

Load package:

```{r, message = FALSE, warning = FALSE}
library(brms)
```

## Generate data by hand

First, generate data by hand:

```{r}
df <- data.frame(x = c(1, 2, 2.5, 3, 4, 5, 6,
                       1, 2, 3.5, 4.5, 5, 5, 6,
                       1, 1, 2.5, 3.5, 4, 5.5, 5.5,
                       1.5, 1.5, 2.5, 3, 4, 4.5, 6),
                 y = c(4.5, 4, 4, 3.5, 2.2, 1.8, 0.9,
                       3, 3, 3.5, 3, 3, 3.5, 2.5,
                       4.5, 3.5, 3.5, 3, 3.5, 2.7, 2.9,
                       2.5, 3, 3, 3, 3, 3.5, 3.5),
                 sub = rep(LETTERS[1:4], each = 7))
```

Next, add a bit more noise and fix some data points (did this after I had checked the model and ran into some convergence issues...):

```{r}
set.seed(42)
df$x <- df$x + rnorm(nrow(df), mean = 0, sd = 0.6)
df$y <- df$y + rnorm(nrow(df), mean = 0, sd = 0.5)
df[4, ]$y <- 2
df[5, ]$y <- 2.5
```

## Fit the models

Fit the models:

```{r, message = FALSE, warning = FALSE}
# Simple linear regression:

brm_lm <- brm(y ~ 1 + x, data = df,
              warmup = 2000, iter = 4000)

# Mixed model:

brm_lmer <- brm(y ~ 1 + x + (1 + x|sub), data = df,
                control = list(adapt_delta = 0.999),
                init = 0, seed = 666,
                warmup = 2000, iter = 3000)
```

Summarize the models:

```{r}
summary(brm_lm)
summary(brm_lmer)
```

Check posterior probabilities of the effect:

```{r}
lm_posts <- posterior_samples(brm_lm)
lmer_posts <- posterior_samples(brm_lmer)

sum(lm_posts$b_x > 0) / nrow(lm_posts)
sum(lmer_posts$b_x > 0) / nrow(lmer_posts)
```


## Plotting

Extract random slopes for plots:

```{r}
ints <- coef(brm_lmer)$sub[, , 1][, 1] # intercepts
slopes <- coef(brm_lmer)$sub[, , 2][, 1] # slopes
```

Create the triple plot and save it as a pdf:

```{r}
pdf(file = '../figures/slope_plot.pdf',
    width = 14, height = 4.5)
# Plot 1:
par(mfrow = c(1, 3), omi = c(0.25, 0.25, 0, 0))
plot(1, 1, type = 'n', xlim = c(0, 7), ylim = c(0, 5.5),
     xaxt = 'n', yaxt = 'n',
     xlab = '', ylab = '')
abline(v = 0, lty = 2)
axis(side = 1, at = 0:7, font = 2)
axis(side = 2, at = 0:5, font = 2, las = 2)
mtext(side = 3, text = '(a) Simple linear regression',
      cex = 1.3, font = 2, line = 1)
mtext(side = 2, text = 'y', cex = 1.5, font = 2, line = 3.5)
mtext(side = 1, text = 'x', cex = 1.5, font = 2, line = 3.6)
points(df$x, df$y, pch = 16, col = rgb(0, 0, 0, 0.5),
       cex = 2)
abline(lm(y ~ x, df), lwd = 3)
points(x = 0, y = as.data.frame(fixef(brm_lm))$Estimate[1],
       pch = 22, bg = 'white', col = 'black', cex = 3)

# Plot 2:
plot(1, 1, type = 'n', xlim = c(0, 7), ylim = c(0, 5.5),
     xaxt = 'n', yaxt = 'n',
     xlab = '', ylab = '')
axis(side = 1, at = 0:7, font = 2)
axis(side = 2, at = 0:5, font = 2, las = 2)
mtext(side = 1, text = 'x', cex = 1.5, font = 2, line = 3.6)
mtext(side = 3, text = '(b) Mixed model with random slopes',
      cex = 1.3, font = 2, line = 1)
points(df$x, df$y, pch = 16, col = rgb(0, 0, 0, 0.5),
       cex = 2)
abline(lm(y ~ x, df), lwd = 3)
for (i in seq_along(ints)) {
  abline(a = ints[i], b = slopes[i], lty = 2)
}

lm_fixef <- as.data.frame(fixef(brm_lm))
lmer_fixef <- as.data.frame(fixef(brm_lmer))

# Plot 3:
plot(1, 1, type = 'n', xlim = c(-1, 1), ylim = c(0, 3),
     xaxt = 'n', yaxt = 'n',
     xlab = '', ylab = '')
axis(side = 1, at = seq(-1, 1, 0.2), font = 2)
abline(v = 0, lty = 2)
text(x = -0.5, y = 2.45, labels = 'without\nrandom slopes',
     cex = 1.6)
points(x = lm_fixef[2, ]$Estimate, y = 2, pch = 15, cex = 2)
arrows(x0 = lm_fixef[2, ]$Q2.5,
       x1 = lm_fixef[2, ]$Q97.5,
       y0 = 2, y1 = 2,
       code = 3, angle = 90, length = 0.1, lwd = 2)
text(x = -0.5, y = 1.45, labels = 'with\nrandom slopes',
     cex = 1.6)
points(x = lmer_fixef[2, ]$Estimate, y = 1, pch = 15, cex = 2)
arrows(x0 = lmer_fixef[2, ]$Q2.5,
       x1 = lmer_fixef[2, ]$Q97.5,
       y0 = 1, y1 = 1,
       code = 3, angle = 90, length = 0.1, lwd = 2)
mtext(side = 3, text = '(c) Effect of x in both models',
      cex = 1.3, font = 2, line = 1)
mtext(side = 1, text = 'Estimated slope', cex = 1.5, font = 2, line = 3.6)
dev.off()
```



