---
title: "Counts of phonetics papers with exact repetitions"
author: "anonymous"
date: "07/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This markdown reports summary counts for exact repetitions (Section 3.7) in the paper, specifically, taken from:

- Roettger and Gordon (2017)
- Nicenboim et al. (2018)
- Winter (2015)
- 2019 issues in JPhon and Language & Speech

## Load data

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

Load the data:

```{r, message = FALSE, warning = FALSE}
nicenboim <- read_csv('../data/nicenboim_incomplete.csv')
winter <- read_csv('../data/JPHon_2014_analysis.csv')
gordon <- read_csv('../data/Gordon&Roettger_Stress Corpus.csv')
newdata <- read_csv('../data/repetition_check_2019_jphon_language&speech.csv')
```

## Roettger and Gordon (2017)

For this dataset, repetition == 1 is no repetition.

```{r}
gordon <- gordon %>%
  mutate(YesNo = ifelse(repetitions > 1, 'yes', 'no'))
```

Count:

```{r}
gordon %>% 
  count(YesNo) %>% 
  mutate(sum = sum(n))
```

Correlate word count with repetition count:

```{r}
with(gordon, cor(no_words, repetitions,
                 method = 'spearman', use = 'complete.obs'))
```

This is quite worrisome. Researchers seem to trade number of items with number of repetitions.

## Nicenboim et al. (2018)

Counter variable:

```{r}
nicenboim <- nicenboim %>% 
  mutate(YesNo = ifelse(rep > 1, 'yes', 'no'))
```

And count:

```{r}
nicenboim %>%
  count(YesNo) %>% 
  mutate(sum = sum(n))
```

Correlate with number of pairs:

```{r}
with(nicenboim,
     cor(pairs_no, rep, method = 'spearman', use = 'complete.obs'))
```

## JPhon 2014 from Winter (2015)

Get only the experimental studies:

```{r}
winter <- winter %>% 
  mutate(repetitions = as.numeric(repetitions)) %>% 
  filter(!is.na(repetitions))
```

Look at the raw values:

```{r}
winter$repetitions
```

Create a counter variable whether there were repetitions or not:

```{r}
winter <- mutate(winter,
                 YesNo = ifelse(repetitions > 0, 'yes', 'no'))
```

Count:

```{r}
winter %>% count(YesNo)
```

How many unique items for those studies?

```{r}
as.numeric(winter$unique_items)
```

How many have less than five items?

```{r}
sum(as.numeric(winter$unique_items) < 5, na.rm = TRUE)
```

Correlate the two:

```{r}
with(winter, cor(repetitions, as.numeric(unique_items),
                 method = 'spearman', use = 'complete.obs'))
```

## New sample of JPhon and Language and Speech (all 2019 issues)

There is a difference between this data and the others in that we didn't read the whole papers but searched for 'repeat___' and 'repet___'.

```{r}
newdata %>%
  mutate(YesNo = !is.na(N_rep)) %>% 
  summarize(N = sum(YesNo))
```

Out of:

```{r}
nrow(newdata)
```




